Galaxy Zoo: Classifications for Galaxies in HST Legacy Imaging

### 2014-03-07 writeup/gz_hubble_data.tex
small changes
minor edits
edits to description

Data release paper for GZ Hubble

@@

\begin{itemize}
\item AEGIS has1 orbit each (~2200 seconds) in F606W (V band) and F814W filter and has been dithered to 0.03 ''/pixel; the imaging covers ~710 arcmin$^2$.

@@

@@

\begin{figure}
\includegraphics[width=60mm]{example_ferengi2.ps}
%\includegraphics[height=80mm,angle=-90]{TuningFork.ps}
\caption{Examples of an obvious spiral and obvious elliptical at $z=0$ which both look similar when redshifted to $z=0.8$. See below for the description of the method for redshirting.  \label{exampleFERENGI}}
\end{figure}

@@

The selection spanned a variety of galaxy morphologies (as indicated by GZ2 classifications), with a range of surface brightnesses, and also spanned the redshift range of SDSS targets (in $N_z = 4$ bins) in order to be optimised for different target redshifts in HST imaging.

@@

\begin{table*}
\caption{Summary of morphological categories \label{morphologies}}
\begin{tabular}{lllc}
\hline\hline
Morphology & Label &  Selection & $N_{\rm objects}$ \\
 & & & ($N_z \times N_\mu$) \\
\hline
Features & Yes & $p_{\rm features} > 0.8$, $p_{rm odd} < 0.1$ & 12 \\ 
 & Int & $0.3 < p_{\rm smooth} < 0.6$, $p_{\rm odd} < 0.1$ & 12 \\ 
 & No &  $p_{\rm smooth} > 0.8$, $p_{\rm odd} < 0.1$ & 12 \\ 
Merger & No & $p_{\rm features} > 0.8$, $p_{\rm odd < 0.1}$, $p_{\rm merger} < 0.1$ & 12\\
& Int & $p_{\rm odd} > 0.5$, $0.1< p_{\rm merger} < 0.4$ & 12 \\ 
& Yes & $p_{\rm odd} > 0.5$, $p_{\rm merger} > 0.4$ & 12\\
Edge-on & Yes &  $p_{\rm edgeon} > 0.8$, $p_{\rm features} > 0.5$ & 12 \\
& Int & $0.4 < p_{\rm edgeon} < 0.8$ , $p_{\rm features} > 0.5$ & 12 \\
& No & $p_{\rm edgeon} < 0.2$, $p_{\rm features} > 0.5$ & 12 \\
Bar & No & $p_{\rm bar} < 0.1$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
& Int & $0.2 < p_{\rm bar} < 0.4$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
& Yes&  $p_{\rm bar} > 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
Visible spiral & No & $p_{\rm spiral} < 0.2$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
& Int &  $0.2 < p_{\rm spiral} < 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
& Yes &  $p_{\rm spiral} > 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
Oblique Bulge Size & No & $p_{\rm nobulge>0.6}$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
& Int &  $p_{\rm justnoticeable}>0.6$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
& Yes & $p_{\rm obvious/dominent}>0.5$\footnote{Edmond does this means $p_{\rm obv}+p_{\rm dom}>0.5$ or $p_{\rm ob} OR p_{\rm dom} >0.5$? }, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
Edge-on bulge shape 
& Round & $p_{\rm rounded}>0.5$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.5$ & 12\\
& Boxy & $p_{\rm boxy}>0.4$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.2$ & 12\\
& Non & $p_{\rm nobulge}>0.5$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.5$ & 12\\
\hline\hline
\end{tabular}
\end{table*}

For each of the four ``target redshifts" ($z = 0.3, 0.5, 0.8$ and 1.0), the images were redshifted in $\Delta z = 0.1$ bins up to $z=1.0$. In addition different evolution models were assumed in FERENGI. This evolution is a crude mechanism that is suppose to mimic the brightness increase of galaxies with increasing redshift. The evolution mechanism allows us to make galaxies brighter as linear function of redshift, and that is it. It is an empirical addition to the magnitude of a galaxy of the form $M' = e z + M$, where $M'$ is the corrected magnitude, $e$ is the evolutionary correction in magnitudes.  (i.e. $e=-1$ essentially brightens the galaxy by 1 magnitude by $z=1$). We ran FERENGI for values of $e$ starting from $e=0$ and increasing up to $e=3.5$ in $\Delta e = 0.5$ steps. Depending on the target, or starting redshift we ran FERENGI for several final redshifts and different evolution models as summarized in Table \ref{ferengi}.

\begin{table}
\caption{Summary of FERENGI runs \label{ferengi}}
\begin{tabular}{lcccccc}
\hline\hline
Starting redshift &  $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm runs}$ & $N_{\rm total}$\\
\hline
0.3 & 8 & 7 & 3.5 & 56 & 4031 \\
0.5 & 6 & 4 & 2.0 & 24 & 1728 \\
0.8 & 3 & 3 & 1.5 & 9 & 648 \\
1.0 & 1 & 3 & 1.5 & 3 & 216 \\
\hline\hline
\end{tabular}
\end{table}

Two series of example redshifted images are shown in Figure \ref{exampleFERENGI}.

@@

@@

@@

>>>>

This will be the data release paper for GZ Hubble. We present the classifications, the methodology for data reduction and corrections for redshift dependent biases in the observed morphologies.

@@

\begin{itemize}
\item AEGIS has 1 orbit each (~2200 seconds) in F606W (V band) and F814W filter and has been dithered to 0.03 ''/pixel; the imaging covers ~710 arcmin$^2$.

@@

\begin{table*}
\caption{Summary of GZ Hubble Samples \label{ferengi}}
\begin{tabular}{lclccc}
\hline\hline
Survey &  $t_{\rm exp}$ & Filters & Resolution & Area & $N_{\rm galaxies}$ \\
 & seconds & & \arcsec/pix & sq arcmin \\
\hline
AEGIS & 2200 & F606W (V) and F814W (I) & 0.03 & 710 \\
COSMOS & 2028 & F814W (I) & 0.05 & 6480\\
GEMS & 2160/2286 & F606W (V) and F850LP (z) & 0.03 & 800 \\
GOODS & 1000-2100s & F435W (B), F606W (V), F775W (i), F850LP (z) & 0.03 & 320 \\
\hline\hline
\end{tabular}
\end{table*}

@@

%-----------------------------------------------------------------------------------------------------------------------------------
\begin{figure}
\begin{center}
\includegraphics[width=60mm]{example_ferengi2.ps}
%\includegraphics[height=80mm,angle=-90]{TuningFork.ps}
\caption{Examples of an obvious spiral and obvious elliptical at $z=0$ which both look similar when redshifted to $z=0.8$. See below for the description of the method for redshirting.  \label{exampleFERENGI}}
\end{center}
\end{figure}
 %-----------------------------------------------------------------------------------------------------------------------------------

@@

The selection spanned a variety of galaxy morphologies (as indicated by GZ2 classifications), with a range of surface brightnesses, and also spanned the redshift range of SDSS targets (in $N_z = 4$ bins) in order to be optimised for different target minimum redshifts in HST imaging.

@@

\begin{table*}
\caption{Summary of morphological categories selected for FERENGI sample \label{morphologies}}
\begin{tabular}{lllc}
\hline\hline
Morphology & Label &  Selection & $N_{\rm objects}$ \\
 & & & ($N_z \times N_\mu$) \\
\hline
Features & Yes & $p_{\rm features} > 0.8$, $p_{\rm odd} < 0.1$ & 12 \\ 
 & Int & $0.3 < p_{\rm smooth} < 0.6$, $p_{\rm odd} < 0.1$ & 12 \\ 
 & No &  $p_{\rm smooth} > 0.8$, $p_{\rm odd} < 0.1$ & 12 \\ 
Merger & No & $p_{\rm features} > 0.8$, $p_{\rm odd < 0.1}$, $p_{\rm merger} < 0.1$ & 12\\
& Int & $p_{\rm odd} > 0.5$, $0.1< p_{\rm merger} < 0.4$ & 12 \\ 
& Yes & $p_{\rm odd} > 0.5$, $p_{\rm merger} > 0.4$ & 12\\
Edge-on & Yes &  $p_{\rm edgeon} > 0.8$, $p_{\rm features} > 0.5$ & 12 \\
& Int & $0.4 < p_{\rm edgeon} < 0.8$ , $p_{\rm features} > 0.5$ & 12 \\
& No & $p_{\rm edgeon} < 0.2$, $p_{\rm features} > 0.5$ & 12 \\
Bar & No & $p_{\rm bar} < 0.1$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
& Int & $0.2 < p_{\rm bar} < 0.4$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
& Yes&  $p_{\rm bar} > 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$ & 24 \\
Visible spiral & No & $p_{\rm spiral} < 0.2$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
& Int &  $0.2 < p_{\rm spiral} < 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
& Yes &  $p_{\rm spiral} > 0.8$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.2$, $p_{\rm bar} < 0.1$ & 12 \\
Oblique Bulge Size & No & $p_{\rm nobulge>0.6}$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
& Int &  $p_{\rm justnoticeable}>0.6$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
& Yes & $p_{\rm obvious/dominent}>0.5$\footnote{Edmond does this means $p_{\rm obv}+p_{\rm dom}>0.5$ or $p_{\rm ob} OR p_{\rm dom} >0.5$? }, $p_{\rm features} > 0.5$, $p_{\rm edgeon} < 0.5$, $p_{\rm bar} < 0.2$ & 12 \\
Edge-on bulge shape 
& Round & $p_{\rm rounded}>0.5$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.5$ & 12\\
& Boxy & $p_{\rm boxy}>0.4$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.2$ & 12\\
& Non & $p_{\rm nobulge}>0.5$, $p_{\rm features} > 0.5$, $p_{\rm edgeon} > 0.5$ & 12\\
\hline\hline
\end{tabular}
\end{table*}

For each of the four ``target redshifts" ($z = 0.3, 0.5, 0.8$ and 1.0), the images were redshifted in $\Delta z = 0.1$ bins up to $z=1.0$. In addition different evolution models were assumed in FERENGI. This evolution is a crude mechanism that is suppose to mimic the brightness increase of galaxies with increasing redshift. The evolution mechanism allows us to make galaxies brighter as linear function of redshift, and that is it. It is an empirical addition to the magnitude of a galaxy of the form $M' = e z + M$, where $M'$ is the corrected magnitude, $e$ is the evolutionary correction in magnitudes.  (i.e. $e=-1$ essentially brightens the galaxy by 1 magnitude by $z=1$). We ran FERENGI for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in $\Delta e = 0.5$ steps. Depending on the target, or starting redshift we ran FERENGI for several final redshifts and different evolution models as summarized in Table \ref{ferengi}.

\begin{table}
\caption{Summary of FERENGI runs \label{ferengi}}
\begin{tabular}{lcccccc}
\hline\hline
$z_{\rm target}$ &  $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm runs}$ & $N_{\rm total}$\\
\hline
0.3 & 8 & 7 & -3.0 & 56 & 4031 \\
0.5 & 6 & 4 & -1.5 & 24 & 1728 \\
0.8 & 3 & 3 & -1.0 & 9 & 648 \\
1.0 & 1 & 3 & -1.0 & 3 & 216 \\
\hline\hline
\end{tabular}
\end{table}

Two series of example redshifted images are shown in Figure \ref{exampleFERENGI}, both for runs with no evolution ($e=0$).

@@

\subsubsection{TODO LIST}
We need to do: 
\begin{itemize}
\item Calculate the magnitudes, surface brightnesses and sizes of the galaxies in the FERENGI images....
\item Plot of magnitude distribution of galaxies in each of the four GZH subsamples with the magnitudes of our fake galaxies over plotted. 
\item Instructions of how to link the $z=0$ $p_X$ values for galaxies with a given size, magnitude (surface brightness) in the GZH images. 
\end{itemize}

@@

Now people go and do science with these awesome Galaxy Zoo Hubble classifications.

@@

KS, KLM, BS, CL thank ASIAA for hosting the ``Citizen Science in Astronomy" workshop at which some of this analysis was done.
